---
title: "State Year 8 Inventory Snapshot"
author: "Ian Combs"
date: "2024-07-15"
output: html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---
```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
knitr::opts_knit$set(root.dir = "../data")
options(width = 88)
library(magrittr)
```





### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

<!-- this is where the DOI would go  [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3675991.svg)](https://doi.org/10.5281/zenodo.3675991)
-->


#### [GitHub repository](https://github.com/icombs2017/nfwfEdrMonitoring){target="_blank"}
###

***
This is the working analysis pipeline to analyze data generated from outplant monitoring for the State Year 8 grant.

***

### All analyses performed with R version `r getRversion()`


# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 


```{r,packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "googlesheets4", "dplyr", "officer","reshape2", "stringr", "flextable", "gridExtra", "ggpubr", "Rmisc", "rcompanion", "RColorBrewer", "googledrive", "gdata", "readxl", "DescTools","patchwork", "FSA", "rstatix", "tidyverse", "lme4", 'PMCRMplus', "EnvStats", "emmeans", "MuMIn", "sjstats", "lmerTest", "gargle", "FSA", "vegan", "gtools", "lubridate", "data.table", "forcats", "paletteer")

```


# Importing *Ex Situ* Data
***
##

# Here we are importing data for our plug counts that were downloaded from Seafoundry 

```{r, ex situ data loading, include = TRUE}



total1 <- read.csv("../data/wholeInventory.csv")


```



*Ex Situ* Inventory (Total)

```{r, ex situ inventory total, include = TRUE}


allSitesTotal1 <- total1 %>% 
  select(species, genotype, site, qty, rack, size) %>% 
  separate(genotype, into = c("speciesCode", "geno"), sep = 2, remove = FALSE) %>% 
  mutate(geno = as.character(geno)) %>% 
  # Removing all non-numeric genotypes (e.g. anything weird like 0?000? etc)
  filter(grepl("^\\d+$", geno)) %>% 
  mutate(across(c(site, geno, genotype, species, speciesCode, rack, size), as.factor)) %>% 
  mutate(
    species = if_else(
      species %in% c("AP", "CN", "DL", "MC", "OF", "PC", "PS"),
      species,  # Keep the original value if it's in the list
      "other"   # Replace with 'other' if it's not
    )
  )

allSitesTotal2 <- allSitesTotal1 %>%
  group_by(species, site, genotype, size) %>%
  dplyr::summarize(total_qty = sum(qty, na.rm = TRUE), .groups = "drop") %>% 
  mutate(size = as.factor(size)) %>% 
  mutate(size = recode(size, "NA" = "TBD", "0" = "TBD"))

allSitesTotal3 <- allSitesTotal2 %>% 
  group_by(species, site) %>% 
  mutate(total_genotypes = n_distinct(genotype)) %>% 
  ungroup %>% 
  pivot_wider(names_from = size, values_from = total_qty, values_fill = list(total_qty = 0)) %>% 
  mutate(totalQty = `TBD` + `0.25` + `0.5` + `0.75` + `1`)

# Correct summing of totals after pivoting
allSitesTotal4 <- allSitesTotal3 %>%
  group_by(species, site) %>%
  dplyr::summarize(
    total_genotypes = first(total_genotypes),  # Keep the total_genotypes column
    total_TBD = sum(`TBD`, na.rm = TRUE),
    total_0_25 = sum(`0.25`, na.rm = TRUE),
    total_1 = sum(`1`, na.rm = TRUE),
    total_0_5 = sum(`0.5`, na.rm = TRUE),
    total_0_75 = sum(`0.75`, na.rm = TRUE),
    totalQty = sum(totalQty, na.rm = TRUE),
  )

allSitesTotal5 <- allSitesTotal4 %>% 
  filter(!str_ends(site, "\\[In Situ\\]"))

allSitesTotalTab <- allSitesTotal5 %>% 
  select(site, species,total_genotypes, total_TBD, total_0_25, total_0_5, total_0_75, total_1, totalQty) %>%
  arrange(site) %>% 
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "species", value = as_paragraph("Species")) %>%
  flextable::compose(part = "header", j = "total_genotypes", value = as_paragraph("# Putative Genotypes")) %>%
  flextable::compose(part = "header", j = "site", value = as_paragraph("Site")) %>%
  flextable::compose(part = "header", j = "total_TBD", value = as_paragraph("To Be Recorded")) %>% 
  flextable::compose(part = "header", j = "total_0_25", value = as_paragraph("1/4")) %>%
  flextable::compose(part = "header", j = "total_0_5", value = as_paragraph("1/2")) %>%
  flextable::compose(part = "header", j = "total_0_75", value = as_paragraph("3/4")) %>%
  flextable::compose(part = "header", j = "total_1", value = as_paragraph("1")) %>%
  flextable::compose(part = "header", j = "totalQty", value = as_paragraph("Total")) %>%
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", j = "totalQty") %>% 
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  add_footer_row(values = c("Total", "", 
          sum(allSitesTotal5$total_genotypes, na.rm = TRUE), 
          sum(allSitesTotal5$total_TBD, na.rm = TRUE), 
          sum(allSitesTotal5$total_0_25, na.rm = TRUE), 
          sum(allSitesTotal5$total_0_5, na.rm = TRUE), 
          sum(allSitesTotal5$total_0_75, na.rm = TRUE), 
          sum(allSitesTotal5$total_1, na.rm = TRUE), 
          sum(allSitesTotal5$totalQty, na.rm = TRUE)),
          colwidths = c(1, 1, 1, 1, 1, 1, 1, 1, 1))

# Print or view the final flextable
allSitesTotalTab


allSitesTotalDoc = read_docx()
allSitesTotalDoc = body_add_flextable(allSitesTotalDoc, value = allSitesTotalTab)
# print(allSitesTotalDoc, target = "../tables/totalInventory.docx")
allSitesTotalTab

```

# Importing *In Situ* Data
***
##

# Here we are importing data for our tree counts that were downloaded from Seafoundry 

```{r, in situ data loading, include = TRUE}



inSitu1 <- readxl::read_xlsx("../data/inSituInventory.xlsx")


```



*In Situ* Inventory (ACER)

```{r, in situ inventory acer, include = TRUE}

inSituACER <- inSitu1 %>% 
  filter(Species %in% "ACER") %>% 
  arrange(Site) 
inSituACERtab <- inSituACER %>% 
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Species", value = as_paragraph("Species")) %>%
  flextable::compose(part = "header", j = "# of Putative Genotypes", value = as_paragraph("# of Putative Genotypes")) %>%
  flextable::compose(part = "header", j = "Site", value = as_paragraph("Site")) %>%
  flextable::compose(part = "header", j = "Total", value = as_paragraph("Total")) %>%
  flextable::compose(part = "header", j = "TLE", value = as_paragraph("TLE")) %>%
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", j = "Total") %>% 
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  add_footer_row(values = c("Total", "", 
          sum(inSituACER$`# of Putative Genotypes`, na.rm = TRUE), 
          sum(inSituACER$Total, na.rm = TRUE), 
          sum(inSituACER$TLE, na.rm = TRUE)),
          colwidths = c(1, 1, 1,1,1))

# Print or view the final flextable
inSituACERtab


inSituACERDoc = read_docx()
inSituACERDoc = body_add_flextable(inSituACERDoc, value = inSituACERtab)
# print(inSituACERDoc, target = "../tables/inSituACERinventory.docx")
inSituACERtab

```

*In Situ* Inventory (Massives)

```{r, in situ inventory massives, include = TRUE}

inSituMSVS <- inSitu1 %>% 
  filter(Species != "ACER", Total > 0) %>%
  select(-c(TLE)) %>% 
  arrange(Site)



inSituMSVStab <- inSituMSVS %>% 
   flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Species", value = as_paragraph("Species")) %>%
  flextable::compose(part = "header", j = "# of Putative Genotypes", value = as_paragraph("# of Putative Genotypes")) %>%
  flextable::compose(part = "header", j = "Site", value = as_paragraph("Site")) %>%
  flextable::compose(part = "header", j = "Total", value = as_paragraph("Total")) %>%
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", j = "Total") %>% 
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  add_footer_row(values = c("Total", "", 
          sum(inSituMSVS$`# of Putative Genotypes`, na.rm = TRUE), 
          sum(inSituMSVS$Total, na.rm = TRUE)),
          colwidths = c(1, 1, 1, 1))

# Print or view the final flextable
inSituMSVStab


inSituMSVSDoc = read_docx()
inSituMSVSDoc = body_add_flextable(inSituMSVSDoc, value = inSituMSVStab)
# print(inSituMSVSDoc, target = "../tables/inSituMSVSinventory.docx")
inSituMSVStab

```

# Production
***
We are going to explore our production numbers during this reporting period. 

## Importing Data
We are importing data taken from Seafoundry that has all of the deltas from our reporting period. We then will remove everything that isn't listed as 'fragging' (e.g. transfers) and just focus on our fragging production. 
```{r, ex situ production, include = TRUE}

frag1 <- read.csv("../data/inventoryDeltas.csv")

frag2 <- frag1 %>% 
    mutate(
    date = as.POSIXct(date, format = "%m/%d/%Y %I:%M:%S %p"),
    # Format as YYYY-MM-DD
    monthYear = paste(format(date, "%Y"), format(date, "%m"), "01", sep = "-"), 
    monthYear = as.POSIXct(monthYear, format = "%Y-%m-%d"),
    # Convert to POSIXct
    monthYear = format(monthYear, "%m/%Y")) %>% 
  filter(monthYear %in% c("10/2024")) %>% 
  select(site, species, genotype_name, reason, qty, comment) %>% 
  mutate(across(c(site, species, genotype_name, reason), as.factor)) %>% 
  filter(reason %in% ("fragging")) %>%
  # mutate(
  #   species = if_else(
  #     species %in% c("AP", "CN", "DL", "MC", "OF", "PC", "PS"),
  #     species,  # Keep the original value if it's in the list
  #     "other"   # Replace with 'other' if it's not
  #   )
  # ) %>% 
  group_by(site, species) %>%
  dplyr::summarize(production  = sum(qty),
                   numberOfGenets = n_distinct(genotype_name)) %>% 
  select(site, species, numberOfGenets, production)



```
##Fragging Table
Making a table to summarize our production Outputs

```{r}

fragTab <- frag2 %>% 
flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "site", value = as_paragraph("Site")) %>%
  flextable::compose(part = "header", j = "species", value = as_paragraph("Species")) %>%
  flextable::compose(part = "header", j = "production", value = as_paragraph("Total Fragments Produced")) %>%
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", j = "production") %>% 
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  add_footer_row(values = c("Total", "", 
          
          sum(frag2$numberOfGenets, na.rm = TRUE),
          sum(frag2$`production`, na.rm = TRUE)),
          colwidths = c(1, 1, 1, 1))
fragTab


fragDoc = read_docx()
fragDoc = body_add_flextable(fragDoc, value = fragTab)
# print(fragDoc, target = "../tables/productionTable.docx")
fragTab



```
 ## Creating a figure to show production numbers
 
 Re-wrangling the data, we are just going to use the month of October from this reporting period (07/01/2024 - 12/31/2024) to allocate toward state year 8 deliverables. 
 
```{r, production figure wrangling, include = TRUE}



frag3 <- frag1 %>% 
 mutate(
    date = as.POSIXct(date, format = "%m/%d/%Y %I:%M:%S %p"),
    # Format as YYYY-MM-DD
    monthYear = paste(format(date, "%Y"), format(date, "%m"), "01", sep = "-"), 
    monthYear = as.POSIXct(monthYear, format = "%Y-%m-%d"),
    # Convert to POSIXct
    monthYear = format(monthYear, "%m/%Y")) %>% 
  select(monthYear, site, species, genotype_name, qty, reason) %>% 
  mutate(across(c(site, species, genotype_name, reason), as.factor)) %>% 
  filter(reason %in% ('fragging')) %>% 
 # mutate(
    # species = if_else(
    #   species %in% c("AP", "CN", "DL", "MC", "OF", "PC", "PS"),
    #   species,  # Keep the original value if it's in the list
    #   "other"   # Replace with 'other' if it's not
    # )) %>% 
  mutate(species = as.factor(species)) %>% 
    filter(monthYear %in% c("10/2024")) %>%
  group_by(monthYear, site, species) %>%
  dplyr::summarize(production  = sum(qty)) %>% 
  mutate(species = case_when(
    species == "INTERSEPTA" ~ "SI",
    species == "RADIANS" ~ "SR",
    species == "SIDEREA" ~ "SI",
    species == "ASTREOIDES" ~ "PA",
    TRUE ~ species  # Keep other values as is
  ))
  
  
  
  
```

Now we are visualizing the data as stacked bar graph with `date` on the X-axis, `production` numbers on the y axis, and stacked by `species`.

```{r, visualizing, include = TRUE}

myColors = paletteer::paletteer_c("ggthemes::Classic Blue", n = 12)


# frag3$species <- factor(frag3$species, levels = frag3$species[order(frag3$production)])

agg_prod <- aggregate(production ~ species, data = frag3, sum)

frag3$species <- factor(frag3$species, levels = agg_prod$species[order(agg_prod$production)])


fragPlot1 <- ggplot(frag3, aes(x = site, y = production, fill = species)) +
  geom_bar(stat = "identity") + # Creates the bar chart
  ylim(0, 4000)+
  scale_fill_manual(values = myColors,
                    labels = c("AP" = "APAL",
               "CN" = "CNAT",
               "MC" = "MCAV",
               "SI" = "SINT",
               "SR" = "SRAD",
               "SS" = "SSID",
               "DL" = "DLAB",
               "OF" = "OFAV",
               "PS" = "PSTR",
               "DS" = "DSTO",
               "PC" = "PCLI", 
               "PA" = "PAST"        # Custom legend labels
  ))+
  labs(
    x = "Site", 
    y = "Production", 
    fill = "Species",
    title = "Coral Fragment Production Through the Reporting Period"
  )

fragPlot <- fragPlot1 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(size = 20),  
    axis.title.x = element_text(size = 40, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 100, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 30, face = "bold"),
    # Facets <- this will be removed if you don't have facets!
    strip.text = element_text(size = 50, face = 'bold'),
    strip.background = element_rect(fill = "#D1D0D0"),
    # Legend
    legend.title = element_text(size = 40, face = "bold"),
    legend.text = element_text(size = 30, face = "bold", colour = "black"),
    legend.position = "right",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = '#f3f3f3'),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Text
    text = element_text(size=20, color="black"))


ggsave("../figures/productionPlot.png", plot = fragPlot, width = 20, height = 15, units = 'in', dpi = 600)



```






