---
title: "State Year 8 Inventory Snapshot"
author: "Ian Combs"
date: "2024-07-15"
output: html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---
```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
knitr::opts_knit$set(root.dir = "../data")
options(width = 88)
library(magrittr)
```





### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

<!-- this is where the DOI would go  [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3675991.svg)](https://doi.org/10.5281/zenodo.3675991)
-->


#### [GitHub repository](https://github.com/icombs2017/nfwfEdrMonitoring){target="_blank"}
###

***
This is the working analysis pipeline to analyze data generated from outplant monitoring for the State Year 8 grant.

***

### All analyses performed with R version `r getRversion()`


# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 


```{r,packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "googlesheets4", "dplyr", "officer","reshape2", "stringr", "flextable", "gridExtra", "ggpubr", "Rmisc", "rcompanion", "RColorBrewer", "googledrive", "gdata", "readxl", "DescTools","patchwork", "FSA", "rstatix", "tidyverse", "lme4", 'PMCRMplus', "EnvStats", "emmeans", "MuMIn", "sjstats", "lmerTest", "gargle", "FSA", "vegan", "gtools", "lubridate", "data.table", "forcats")

```


# Importing *Ex Situ* Data
***
##

# Here we are importing data for our plug counts that were downloaded from Seafoundry 

```{r, ex situ data loading, include = TRUE}



total1 <- read.csv("../data/wholeInventory.csv")


```



*Ex Situ* Inventory (Total)

```{r, ex situ inventory total, include = TRUE}


allSitesTotal1 <- total1 %>% 
  select(species, genotype, site, qty, rack, size) %>% 
  separate(genotype, into = c("speciesCode", "geno"), sep = 2, remove = FALSE) %>% 
  mutate(geno = as.character(geno)) %>% 
  # Removing all non-numeric genotypes (e.g. anything weird like 0?000? etc)
  filter(grepl("^\\d+$", geno)) %>% 
  mutate(across(c(site, geno, genotype, species, speciesCode, rack, size), as.factor)) %>% 
  mutate(
    species = if_else(
      species %in% c("AP", "CN", "DL", "MC", "OF", "PC", "PS"),
      species,  # Keep the original value if it's in the list
      "other"   # Replace with 'other' if it's not
    )
  )

allSitesTotal2 <- allSitesTotal1 %>%
  group_by(species, site, genotype, size) %>%
  dplyr::summarize(total_qty = sum(qty, na.rm = TRUE), .groups = "drop") %>% 
  mutate(size = as.factor(size)) %>% 
  mutate(size = recode(size, "NA" = "TBD", "0" = "TBD"))

allSitesTotal3 <- allSitesTotal2 %>% 
  group_by(species, site) %>% 
  mutate(total_genotypes = n_distinct(genotype)) %>% 
  ungroup %>% 
  pivot_wider(names_from = size, values_from = total_qty, values_fill = list(total_qty = 0)) %>% 
  mutate(totalQty = `TBD` + `0.25` + `0.5` + `0.75` + `1`)

# Correct summing of totals after pivoting
allSitesTotal4 <- allSitesTotal3 %>%
  group_by(species, site) %>%
  dplyr::summarize(
    total_genotypes = first(total_genotypes),  # Keep the total_genotypes column
    total_TBD = sum(`TBD`, na.rm = TRUE),
    total_0_25 = sum(`0.25`, na.rm = TRUE),
    total_1 = sum(`1`, na.rm = TRUE),
    total_0_5 = sum(`0.5`, na.rm = TRUE),
    total_0_75 = sum(`0.75`, na.rm = TRUE),
    totalQty = sum(totalQty, na.rm = TRUE),
  )

allSitesTotal5 <- allSitesTotal4 %>% 
  filter(!str_ends(site, "\\[In Situ\\]"))

allSitesTotalTab <- allSitesTotal5 %>% 
  select(site, species,total_genotypes, total_TBD, total_0_25, total_0_5, total_0_75, total_1, totalQty) %>%
  arrange(site) %>% 
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "species", value = as_paragraph("Species")) %>%
  flextable::compose(part = "header", j = "total_genotypes", value = as_paragraph("# Putative Genotypes")) %>%
  flextable::compose(part = "header", j = "site", value = as_paragraph("Site")) %>%
  flextable::compose(part = "header", j = "total_TBD", value = as_paragraph("To Be Recorded")) %>% 
  flextable::compose(part = "header", j = "total_0_25", value = as_paragraph("1/4")) %>%
  flextable::compose(part = "header", j = "total_0_5", value = as_paragraph("1/2")) %>%
  flextable::compose(part = "header", j = "total_0_75", value = as_paragraph("3/4")) %>%
  flextable::compose(part = "header", j = "total_1", value = as_paragraph("1")) %>%
  flextable::compose(part = "header", j = "totalQty", value = as_paragraph("Total")) %>%
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", j = "totalQty") %>% 
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  add_footer_row(values = c("Total", "", 
          sum(allSitesTotal5$total_genotypes, na.rm = TRUE), 
          sum(allSitesTotal5$total_TBD, na.rm = TRUE), 
          sum(allSitesTotal5$total_0_25, na.rm = TRUE), 
          sum(allSitesTotal5$total_0_5, na.rm = TRUE), 
          sum(allSitesTotal5$total_0_75, na.rm = TRUE), 
          sum(allSitesTotal5$total_1, na.rm = TRUE), 
          sum(allSitesTotal5$totalQty, na.rm = TRUE)),
          colwidths = c(1, 1, 1, 1, 1, 1, 1, 1, 1))

# Print or view the final flextable
allSitesTotalTab


allSitesTotalDoc = read_docx()
allSitesTotalDoc = body_add_flextable(allSitesTotalDoc, value = allSitesTotalTab)
# print(allSitesTotalDoc, target = "../tables/totalInventory.docx")
allSitesTotalTab

```

# Importing *In Situ* Data
***
##

# Here we are importing data for our tree counts that were downloaded from Seafoundry 

```{r, in situ data loading, include = TRUE}



inSitu1 <- readxl::read_xlsx("../data/inSituInventory.xlsx")


```



*In Situ* Inventory (ACER)

```{r, in situ inventory acer, include = TRUE}


inSituACERtab <- inSitu1 %>% 
  filter(Species %in% "ACER") %>% 
  arrange(Site) %>% 
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Species", value = as_paragraph("Species")) %>%
  flextable::compose(part = "header", j = "# of Putative Genotypes", value = as_paragraph("# of Putative Genotypes")) %>%
  flextable::compose(part = "header", j = "Site", value = as_paragraph("Site")) %>%
  flextable::compose(part = "header", j = "Total", value = as_paragraph("Total")) %>%
  flextable::compose(part = "header", j = "TLE", value = as_paragraph("TLE")) %>%
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", j = "Total") %>% 
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  add_footer_row(values = c("Total", "", 
          sum(inSitu1$`# of Putative Genotypes`, na.rm = TRUE), 
          sum(inSitu1$Total, na.rm = TRUE), 
          sum(inSitu1$TLE, na.rm = TRUE)),
          colwidths = c(1, 1, 1,1,1))

# Print or view the final flextable
inSituACERtab


inSituACERDoc = read_docx()
inSituACERDoc = body_add_flextable(inSituACERDoc, value = inSituACERtab)
# print(inSituACERDoc, target = "../tables/inSituACERinventory.docx")
inSituACERtab

```

*In Situ* Inventory (Massives)

```{r, in situ inventory massives, include = TRUE}


inSituMSVStab <- inSitu1 %>% 
  filter(Species != "ACER", Total > 0) %>%
  select(-c(TLE)) %>% 
  arrange(Site) %>% 
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Species", value = as_paragraph("Species")) %>%
  flextable::compose(part = "header", j = "# of Putative Genotypes", value = as_paragraph("# of Putative Genotypes")) %>%
  flextable::compose(part = "header", j = "Site", value = as_paragraph("Site")) %>%
  flextable::compose(part = "header", j = "Total", value = as_paragraph("Total")) %>%
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", j = "Total") %>% 
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  add_footer_row(values = c("Total", "", 
          sum(inSitu1$`# of Putative Genotypes`, na.rm = TRUE), 
          sum(inSitu1$Total, na.rm = TRUE)),
          colwidths = c(1, 1, 1, 1))

# Print or view the final flextable
inSituMSVStab


inSituMSVSDoc = read_docx()
inSituMSVSDoc = body_add_flextable(inSituMSVSDoc, value = inSituMSVStab)
# print(inSituMSVSDoc, target = "../tables/inSituMSVSinventory.docx")
inSituMSVStab

```



