---
title: "State Year 7 Report"
author: "Ian Combs"
date: "2024-1-08"
output: html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---
```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
knitr::opts_knit$set(root.dir = "../data")
options(width = 88)
library(magrittr)
```





### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

<!-- this is where the DOI would go  [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3675991.svg)](https://doi.org/10.5281/zenodo.3675991)
-->


#### [GitHub repository](https://github.com/icombs2017/nfwfEdrMonitoring){target="_blank"}
###

***
This is the working analysis pipeline to analyze data generated from outplant monitoring for the State Year 7 grant.

***

### All analyses performed with R version `r getRversion()`


# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 


```{r,packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "googlesheets4", "dplyr", "officer","reshape2", "stringr", "flextable", "gridExtra", "ggpubr", "Rmisc", "rcompanion", "RColorBrewer", "googledrive", "gdata", "readxl", "DescTools","patchwork", "FSA", "rstatix", "tidyverse", "lme4", 'PMCRMplus', "EnvStats", "emmeans", "MuMIn", "sjstats", "lmerTest", "gargle", "FSA", "vegan", "gtools", "lubridate", "data.table", "forcats")

```


# Importing Data
***
## We are downloading this dataset from our GoogleDrive folder. We will be using the package `googledrive`. Each GoogleDrive file has a unique ID that does not change throughout the lifespan of the document, even if the file name is changed. This ID is housed in the file's URL. 
Example: docs.google.com/spreadsheets/d/FILE_ID_GOES_HERE/other_information/. Below you will copy and paste that into the function `drive_download` within `as_id`. This will save the file locally in the specified path (in this case our data folder) and you will import the folder as you normally would. Downloading it this way decreases the potential human error when downloading, moving folders, renaming, saving etc and ensures that the most up to date file is being utilized. 

# Here we are importing data for our plug counts from google drive

```{r, outplantLoadingData, include = TRUE}

data <- drive_download(
 as_id("18w_T068tjY1kse9qKBL1U_gDjZ9iByMcpmUIS5R4tk8"),
   path = "../data/outplantData.xlsx",
   overwrite = TRUE)

# The start and end dates for this reporting period
reportingStart <- as.Date("2024-07-01")
reportingEnd <- as.Date("2025-06-30")

#Only Site with 90day data

y8Data <- read_excel("../data/outplantData.xlsx", sheet = 1, skip = 1, col_names = FALSE)

# Step 2: Set the first row as column names
colnames(y8Data) <- as.character(y8Data[1, ])  # use row 1 as header
y8Data <- y8Data[-1, ]  # remove the first row now that it's the header


outplantRaw <- read.csv("../data/outplanting.csv")


```


# 2024 Monitoring

## Time to do some data wrangling for the State Year 7 dataset. Since we only care about the State Year 7 Deliverables right now, we are going to filter the data set to just include those given to the grant *State Yr7*. Also including a fixed string called *deliverables* and *cumulativeDeliverables* that includes what each deliverable is each year and cumulatively by year.



```{r, apalDataWrangling, include = TRUE}

outplantY8 <- y8Data %>%
  select(`Outplant Date`,
         `Site ID`, 
         `# of frags`, 
         `90 day # Dead`, 
         `90 day # Missing`
  ) %>% 
  filter(`Site ID` %in% "IC_Z2") %>%
  mutate(site = as.factor(`Site ID`),
         frags = as.numeric(`# of frags`),
         numberDead = as.numeric(`90 day # Dead`),
         numberMissing = as.numeric(`90 day # Missing`)) %>% 
  mutate(percentSurvival = ((frags - (numberDead+numberMissing))/ frags) * 100,
         percentDead = ((numberDead/frags)*100),
        percentMissing = (numberMissing/frags)*100) %>% 
  select(site, frags, numberDead, numberMissing, percentSurvival, percentDead, percentMissing)


outplanting <- outplantRaw %>% filter(Deliverable %in% "State Year 8")

coral1 <- outplanting %>% 
  select(Location, Date, Latitude, Longitude, Genotype, Outplanted) %>% 
  separate(Location, into = c("trash", "site"), sep = "/") %>% 
  filter(!site %in% " Junk Data") %>% 
  droplevels() %>% 
  select(Date, site, Genotype, Latitude, Longitude,Outplanted) %>% 
  separate(Genotype, into = c("speciesCode", "geno"), sep = 2, remove = FALSE) %>%
  mutate(geno = as.character(geno)) %>% 
  mutate(site = as.factor(site),
         Genotype = as.factor(Genotype),
         speciesCode = as.factor(speciesCode),
         geno = as.factor(geno)) %>% 
  mutate( Date = as.Date(Date, "%m/%d/%y")) %>% 
  arrange(Date) %>% 
  group_by(Date, site, Latitude, Longitude, speciesCode, Genotype) %>% 
  dplyr::summarize(totalGenoOutplanted = sum(Outplanted)) %>%
  group_by(Date, site, Latitude, Longitude, speciesCode) %>% 
  dplyr::mutate(totalSpeciesOutplanted = sum(totalGenoOutplanted)) %>% 
group_by(Date, site, Latitude, Longitude) %>% 
  mutate(totalOutplanted = sum(totalGenoOutplanted))
   



deliverables <- c(10000)

cumulativeDeliverables <- c(10000)



```


# Plotting 2024 Outplant Numbers

## Plotting outplant numbers for all 2024 corals.


```{r, 2024OutplantCumulativePlot, include = TRUE}

outplantY24plot1<- ggplot(coral1, aes(x = Date, y = cum_sum))+
  geom_line(aes(color = "Fragments Outplanted", linetype = 'solid'), linewidth = 3)+
  geom_hline(aes(yintercept = cumulativeDeliverables, linetype = 'State Year 7 Outplant Goal', color = 'State Year 7 Outplant Goal'), linewidth = 3)+
  scale_y_continuous(limits = c(0, 15000), breaks = seq(0, 15000, by = 500))+
  ggtitle(expression("Coral Fragments Outplanted for State Year 7"))+
  xlab(" ")+
  ylab("Number of fragments outplanted")+
  scale_color_manual(values = c('blue', 'black'), name = " ") +
  scale_linetype_manual(name = " ", values = c('solid', 'dashed', 'dashed', 'dashed'), guide = "none")




outplantY24plot <- outplantY24plot1 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))


ggsave("../figures/cumulativeY24outplantPlot.png", plot = outplantY24plot, width = 20, height = 15, units = 'in', dpi = 600)

```
## 2024 Outplant Table
Creating a table for number of outplant events


```{r, y24OutplantTable, include = TRUE}


# Making the table
outplantY24tab <- coral1 %>% 
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Date", value = as_paragraph("Outplant Date")) %>%
  flextable::compose(part = "header", j = "totalOutplanted", value = as_paragraph("Number of Corals Outplanted")) %>%
  flextable::compose(part = "header", j = "Latitude", value = as_paragraph("Latitude")) %>% 
  flextable::compose(part = "header", j = "Longitude", value = as_paragraph("Longitude"))%>%
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", j = "totalOutplanted") %>% 
  set_table_properties(width = 1, align = "center", layout = "autofit")


outplantY24Doc = read_docx()
outplantY24Doc = body_add_flextable(outplantY24Doc, value = outplantY24tab)
print(outplantY24Doc, target = "../tables/outplantStateY8table.docx")
outplantY24tab








```
## Y24 One Month Survival Plot
Plotting One Month Survival Using a stacked bar graph
```{r, oneMonthSurvivalPlotY24, include = TRUE}

myColors <- c("red3", "yellow3", "green3")


y24OutplantPlotData <- outplantData %>% 
pivot_longer(cols = c("1-month Survival", "1-month % Dead", "1-month % Missing"), names_to = "1 month", values_to = "1 mo percentage") %>% 
  mutate(`1 month` = as.factor(`1 month`))


y241MoSurvivalPlot1<- ggplot(y24OutplantPlotData, aes(x = `Site ID`, y = `1 mo percentage`, fill = `1 month`, color = I('black')))+
  geom_bar(position = "stack", stat = 'identity', width = 1)+
  ggtitle(expression("1 Month Proportional Survival of Coral Outplants"))+
  xlab("Site")+
  ylab("Proportional Survival")+
  scale_fill_manual(values = myColors, labels = c("Dead", "Missing", "Alive"))

y241MoSurvivalPlot <- y241MoSurvivalPlot1 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(size = 30, angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 40, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 40, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 30, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_blank(),
    legend.text = element_text(size = 40, face = "bold", colour = "black"),
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))




ggsave("../figures/oneMonthSurvivalY24.png", plot =y241MoSurvivalPlot, width = 20, height = 15, units = 'in', dpi = 600)



```


everything above might be garbage!

```{r, outplant, include = TRUE}



outplant1 <- read.csv("../data/outplanting.csv")

outplant2 <- outplant1 %>% filter(Deliverable == "State Year 8")





```





```{r}
coral1

genotype <- coral1 %>% 
  





```


