---
title: "State Year 7 Bleach Monitoring"
author: "Ian Combs"
date: "2024-7-15"
output: html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---
```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
knitr::opts_knit$set(root.dir = "../data")
options(width = 88)
library(magrittr)
```





### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

<!-- this is where the DOI would go  [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3675991.svg)](https://doi.org/10.5281/zenodo.3675991)
-->


#### [GitHub repository](https://github.com/icombs2017/nfwfEdrMonitoring){target="_blank"}
###

***
This is the working analysis pipeline to analyze data generated from outplant monitoring for the State Year 7 grant.

***

### All analyses performed with R version `r getRversion()`


# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 


```{r,packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "googlesheets4", "dplyr", "officer","reshape2", "stringr", "flextable", "gridExtra", "ggpubr", "Rmisc", "rcompanion", "RColorBrewer", "googledrive", "gdata", "readxl", "DescTools","patchwork", "FSA", "rstatix", "tidyverse", "lme4", 'PMCRMplus', "EnvStats", "emmeans", "MuMIn", "sjstats", "lmerTest", "gargle", "FSA", "vegan", "gtools", "lubridate", "data.table", "forcats")

```


# Importing Data
***
## We are downloading this dataset from our GoogleDrive folder. We will be using the package `googledrive`. Each GoogleDrive file has a unique ID that does not change throughout the lifespan of the document, even if the file name is changed. This ID is housed in the file's URL. 
Example: docs.google.com/spreadsheets/d/FILE_ID_GOES_HERE/other_information/. Below you will copy and paste that into the function `drive_download` within `as_id`. This will save the file locally in the specified path (in this case our data folder) and you will import the folder as you normally would. Downloading it this way decreases the potential human error when downloading, moving folders, renaming, saving etc and ensures that the most up to date file is being utilized. 

# Here we are importing data for our plug counts from google drive

```{r, outplantLoadingData, include = TRUE}

# data <- drive_download(
#  as_id("1jdOc145IpKYU2hctqFZfJNGsVMGUoj2P"),
#    path = "../data/bleachingData.xlsx",
#    overwrite = TRUE)

bd <- read_excel("../data/bleachingData.xlsx", na = "N/A")


speciesBleachACER <- bd %>%
  mutate(across(c(Deliverable,Location,`Site ID`), as.factor)) %>% 
  select(-c(Apal, Ofav,Pcli,Mcav)) %>% 
  filter(!is.na(Acer)) %>% 
  select(`Outplant Date`, 
         `Site ID`, 
         Deliverable,
         `total # frags outplanted`,
         `2024 post-bleaching monitoring date`,
         `2024 post-bleaching % survival`, 
         `2024 post-bleaching % dead`, 
         `2024 post-bleaching % missing`) %>% 
  mutate_at(c("Site ID", 
        "Deliverable"), as.factor) %>% 
  droplevels() %>% 
  mutate(`Outplant Date` = as.Date(`Outplant Date`, "%m %d %Y"))
  
speciesBleachAPAL <- bd %>%
  mutate(across(c(Deliverable,Location,`Site ID`), as.factor)) %>% 
  select(-c(Acer, Ofav,Pcli,Mcav)) %>% 
  filter(!is.na(Apal)) %>% 
   select(`Outplant Date`, 
         `Site ID`, 
         Deliverable,
        `total # frags outplanted`,
         `2024 post-bleaching monitoring date`,
         `2024 post-bleaching % survival`, 
         `2024 post-bleaching % dead`, 
         `2024 post-bleaching % missing`,) %>% 
  mutate_at(c("Site ID", 
        "Deliverable"), as.factor) %>% 
  droplevels() %>% 
  mutate(`Outplant Date` = as.Date(`Outplant Date`, "%m %d %Y")) 

  
speciesBleachMsvs <- bd %>%
  mutate(across(c(Deliverable,Location,`Site ID`), as.factor)) %>% 
  select(-c(Acer, Apal)) %>% 
  slice_tail(n = 4) %>% 
   select(`Outplant Date`, 
         `Site ID`, 
         Deliverable, 
         `total # frags outplanted`,
         `2024 post-bleaching monitoring date`,
         `2024 post-bleaching % survival`, 
         `2024 post-bleaching % dead`, 
         `2024 post-bleaching % missing`,) %>% 
  mutate_at(c("Site ID", 
        "Deliverable"), as.factor) %>% 
  droplevels() %>% 
  mutate(`Outplant Date` = as.Date(`Outplant Date`, "%m %d %Y"))

speciesBleachMsvs




```


# Bleaching Monitoring summary



```{r, bleachingDataWrangling, include = TRUE}


bleach <- bd %>% 
  select(`Outplant Date`, 
         `Site ID`, 
         Deliverable, 
         `2024 post-bleaching monitoring date`,
         `2024 post-bleaching % survival`, 
         `2024 post-bleaching % dead`, 
         `2024 post-bleaching % missing`,) %>% 
  mutate_at(c("Site ID", 
        "Deliverable"), as.factor) %>% 
  droplevels() %>% 
  mutate(`Outplant Date` = as.Date(`Outplant Date`, "%m %d %Y")) %>% 
  pivot_longer(cols = c("2024 post-bleaching % survival", "2024 post-bleaching % dead", "2024 post-bleaching % missing"), names_to = "1 month", values_to = "1 mo percentage") %>%   select(`Outplant Date`, `Site ID`, Deliverable, `1 month`, `1 mo percentage`) %>% 
  mutate_at(c("1 month"), as.factor) %>% 
  mutate(`Site ID` = factor(`Site ID`, levels = mixedsort(levels(`Site ID`))))



```

# Plotting APAL Outplant Numbers

## Plotting outplant numbers for all sites during the bleaching event over the course of the course of the grant.



## APAL One Month Survival Plot
Plotting One Month Survival Using a stacked bar graph
```{r, oneMonthSurvivalPlotAPAL, include = TRUE}
myColors <- c("red3", "yellow3", "green3")



bleachSurvivalPlot1<- ggplot(bleach, aes(x = `Site ID`, y = `1 mo percentage`, fill = `1 month`, color = I('black')))+
  geom_bar(position = "stack", stat = 'identity', width = 1)+
  ggtitle(expression("Post-Bleaching Event Proportional Survival"))+
  xlab("Site")+
  ylab("Proportional Survival")+
  scale_fill_manual(values = myColors) + 
  facet_wrap(.~ Deliverable, scales = "free_x")
  

bleachSurvivalPlot <- bleachSurvivalPlot1 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 40, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 40, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 30, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_blank(),
    legend.text = element_text(size = 30, face = "bold", colour = "black"),
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"),
    # Text above facets
    strip.text = element_text(size = 20))




ggsave("../figures/postBleachingSurvival.png", plot =bleachSurvivalPlot, width = 30, height = 30, units = 'in', dpi = 600)



```






```{r, summaryStats, include = TRUE}
# by State Year
acerSummary <- speciesBleachACER %>% 
  group_by(Deliverable) %>% 
  dplyr::summarize(meanSurvival = mean((`2024 post-bleaching % survival`)*100),
                   meanDead = mean((`2024 post-bleaching % dead`)*100),
                   meanMissing = mean((`2024 post-bleaching % missing`)*100))

apalSummary <- speciesBleachAPAL %>% 
  group_by(Deliverable) %>% 
  dplyr::summarize(meanSurvival = mean((`2024 post-bleaching % survival`)*100),
                   meanDead = mean((`2024 post-bleaching % dead`)*100),
                   meanMissing = mean((`2024 post-bleaching % missing`)*100))


msvsSummary <- speciesBleachMsvs %>% 
  group_by(Deliverable) %>% 
  dplyr::summarize(meanSurvival = mean((`2024 post-bleaching % survival`)*100),
                   meanDead = mean((`2024 post-bleaching % dead`)*100),
                   meanMissing = mean((`2024 post-bleaching % missing`)*100))

# Overall
acerSummaryOverall <- speciesBleachACER %>% 
  # group_by(Deliverable) %>% 
  dplyr::summarize(meanSurvival = mean((`2024 post-bleaching % survival`)*100),
                   meanDead = mean((`2024 post-bleaching % dead`)*100),
                   meanMissing = mean((`2024 post-bleaching % missing`)*100))

apalSummaryOverall <- speciesBleachAPAL %>% 
  # group_by(Deliverable) %>% 
  dplyr::summarize(meanSurvival = mean((`2024 post-bleaching % survival`)*100),
                   meanDead = mean((`2024 post-bleaching % dead`)*100),
                   meanMissing = mean((`2024 post-bleaching % missing`)*100))


msvsSummaryOverall <- speciesBleachMsvs %>% 
  # group_by(Deliverable) %>% 
  na.omit() %>% 
  dplyr::summarize(meanSurvival = mean((`2024 post-bleaching % survival`)*100),
                   meanDead = mean((`2024 post-bleaching % dead`)*100),
                   meanMissing = mean((`2024 post-bleaching % missing`)*100))

```



```{r}

totalSurvival <- bd %>% 
  ungroup() %>% 
  dplyr::mutate(fragsAlive = `total # frags outplanted`*`2024 post-bleaching % survival`) %>% 
  dplyr::summarize(totalAlive = sum(fragsAlive),
                   totalOutplanted = sum(`total # frags outplanted`),
                   totalSurvival = (totalAlive/(totalOutplanted)*100))
  
  
acerSurvival <- speciesBleachACER %>% 
  ungroup() %>% 
  dplyr::mutate(fragsAlive = `total # frags outplanted`*`2024 post-bleaching % survival`) %>% 
  dplyr::summarize(totalAlive = sum(fragsAlive),
                   totalOutplanted = sum(`total # frags outplanted`),
                   totalSurvival = (totalAlive/(totalOutplanted)*100),
                   Species = "ACER")

apalSurvival <- speciesBleachAPAL %>% 
  ungroup() %>% 
  dplyr::mutate(fragsAlive = `total # frags outplanted`*`2024 post-bleaching % survival`) %>% 
  dplyr::summarize(totalAlive = sum(fragsAlive),
                   totalOutplanted = sum(`total # frags outplanted`),
                   totalSurvival = (totalAlive/(totalOutplanted)*100),
                    Species = "APAL")



  msvsSurvival <- speciesBleachMsvs %>% 
  ungroup() %>% 
  dplyr::mutate(fragsAlive = `total # frags outplanted`*`2024 post-bleaching % survival`) %>% 
  dplyr::summarize(totalAlive = sum(fragsAlive),
                   totalOutplanted = sum(`total # frags outplanted`),
                   totalSurvival = (totalAlive/(totalOutplanted)*100),
                   Species = "Massives")
  
totalSurvival <- rbind(acerSurvival, apalSurvival, msvsSurvival) %>% 
  select(Species,totalAlive,totalOutplanted,totalSurvival)

bleachSurvivalTab <- totalSurvival %>% 
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Species", value = as_paragraph("Outplanted Species")) %>%
  flextable::compose(part = "header", j = "totalAlive", value = as_paragraph("Fragments Alive")) %>%
  flextable::compose(part = "header", j = "totalOutplanted", value = as_paragraph("Corals Outplanted")) %>% 
  flextable::compose(part = "header", j = "totalSurvival", value = as_paragraph("% Survival")) %>% 
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>% 
  colformat_num(j = "totalAlive", digits = 0) %>%      
  colformat_num(j = "totalSurvival", digits = 2)
  # align(align = "center", j = "# plugs") %>% 
  # set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  # add_footer_row(values = 
  #   list(
  #     `Site ID` = "Averages:",
  #     "",
  #     "",
  #     `1-month Survival` = round(mean(oneMonthAPALtab1$`1-month Survival`), 2),
  #     `1-month % Dead` = round(mean(oneMonthAPALtab1$`1-month % Dead`), 2),
  #     `1-month % Missing` = round(mean(oneMonthAPALtab1$`1-month % Missing`), 2),
  #     `1-month % Plot Survival` = round(mean(oneMonthAPALtab1$`1-month % Plot Survival`),2)),
  #   colwidths = c(1, 1, 1, 1, 1, 1, 1))

 


bleachSurvivalDoc = read_docx()
bleachSurvivalDoc = body_add_flextable(bleachSurvivalDoc, value = bleachSurvivalTab)
print(bleachSurvivalDoc, target = "../tables/bleachSurvivalTABLE.docx")
bleachSurvivalTab


```

survivalFromBleaching 

loading data


```{r, survivalFromBleaching, include = TRUE}

dataACRO <- drive_download(
 as_id("1fpsE_0655xguD48cGK6n1ZqSLnI-tw9aHzhbPM9kZ-k"),
   path = "../data/bleachingDataACRO.xlsx",
   overwrite = TRUE)

bdACER <- read_excel("../data/bleachingDataACRO.xlsx", na = c("N/A", "no data"), sheet = "State Monitoring Summary ACER", n_max = 53)
bdAPAL <- read_excel("../data/bleachingDataACRO.xlsx", na = c("N/A", "no data"), sheet = "State Monitoring Summary APAL")

dataMSVS <- drive_download(
 as_id("1Znbr4E6mkO9PFO8q1b9cbcIWseMc-Kw62JkxgDMQp5s"),
   path = "../data/bleachingDataMSVS.xlsx",
   overwrite = TRUE)

bdMSVS <- read_excel("../data/bleachingDataMSVS.xlsx", na = c("N/A", "no data"), sheet = "massive monitoring summary")

```
wranglingData
```{r}

acerBL1mo <- bdACER %>% 
  select(Deliverable,
         Location,
         `Site ID`,
         `# frags`,
         `1-month % Survival`,
         `1-month % dead`,
         `1-month % missing`, 
         `12-month Monitoring Date`,
         `12-month % Survival`,
         `12-month % dead`, 
         `12-month % missing`,               
        `2024 post-bleaching % survival`,     
        `2024 post-bleaching % dead`,
        `2024 post-bleaching % missing`,    
        `2024 post-bleaching # dead`,
        `2024 post-bleaching # missing`) %>% 
  mutate(`12-month % Survival` = as.numeric(`12-month % Survival`),
        across(where(is.character), as.factor)) %>% 
  filter(str_detect(Deliverable, "State")) %>% 
  dplyr::filter(`12-month Monitoring Date` %in% NA)
  

apalBL <- bdAPAL %>%
  select(Deliverable, 
         Location, 
         `Site ID`,
         `# frags`,
         `1-month Monitoring Date`,
         `1-month % Survival`,
         `1-month % dead`,
         `1-month % missing`,
         `12-month Monitoring Date`,
         `12-month % Survival`,
         `# dead...18`,
         `# missing...19`,
         `2024 post-bleaching % survival`,
         `2024 post-bleaching % dead`,
         `2024 post-bleaching % missing`,
         `# dead...24`,
         `# missing...25`) %>%
    mutate(`12-month % Survival` = as.numeric(`12-month % Survival`),
        across(where(is.character), as.factor)) %>% 
  filter(str_detect(Deliverable, "State"))
  

msvsBL <- bdMSVS %>% 
  select(Deliverable,
  Location,                                    
 `Site ID`,
 `# plugs`,
 `1-month % Survival`,
 `12-month Monitoring Date`,
 `12-month % Survival`,                          
 `2023 post-bleaching monitoring date`,
 `2023 rapid post-bleaching monitoring % survival`,
 `Frags alive`) %>% 
  mutate(`12-month % Survival` = as.numeric(`12-month % Survival`),
        `2023 rapid post-bleaching monitoring % survival` = as.numeric(`2023 rapid post-bleaching monitoring % survival`),
        across(where(is.character), as.factor)) %>% 
  filter(str_detect(Deliverable, "State"))










  
```











```{r}
bleachSurvival1mo <- bd %>% 
  dplyr::filter(!`12-month Monitoring Date` %in% NA) %>% 
  dplyr::mutate(aliveBeforeBleach = (`total # frags outplanted`* `1-month % Survival`),
                deadAfterBleach = (`total # frags outplanted` * `2024 post-bleaching % dead`),
                aliveAfterBleach = (`total # frags outplanted` * `2024 post-bleaching % survival`),
                missingAfterBleach = (aliveBeforeBleach - deadAfterBleach+aliveAfterBleach),
                bleachSurvival = (aliveAfterBleach/aliveBeforeBleach),
                bleachDead = (deadAfterBleach/aliveBeforeBleach),
                bleachMissing = (missingAfterBleach/aliveBeforeBleach),
                other = (1-bleachSurvival)) %>% 
  select(`Outplant Date`, 
         `Site ID`, 
         species,
         Deliverable, 
         bleachSurvival,
         other) %>% 
  mutate_at(c("Site ID", 
        "Deliverable", "species"), as.factor) %>% 
  droplevels() %>% 
  mutate(`Outplant Date` = as.Date(`Outplant Date`, "%m %d %Y")) %>% 
  pivot_longer(cols = c("bleachSurvival", "other"), names_to = "monitoringCategory", values_to = "proportion") %>%   
  select(`Outplant Date`, `Site ID`,species, Deliverable, monitoringCategory, proportion) %>%
  mutate_at(c("monitoringCategory"), as.factor) %>%
  mutate(`Site ID` = factor(`Site ID`, levels = mixedsort(levels(`Site ID`))))





bleachSurvival12mo <- bd %>% 
  dplyr::filter(!`12-month Monitoring Date` %in% NA) %>% 
  dplyr::mutate(aliveBeforeBleach = (`total # frags outplanted`* `12-month % Survival`),
                deadAfterBleach = (`total # frags outplanted` * `2024 post-bleaching % dead`),
                aliveAfterBleach = (`total # frags outplanted` * `2024 post-bleaching % survival`),
                missingAfterBleach = (aliveBeforeBleach - deadAfterBleach+aliveAfterBleach),
                bleachSurvival = (aliveAfterBleach/aliveBeforeBleach),
                bleachDead = (deadAfterBleach/aliveBeforeBleach),
                bleachMissing = (missingAfterBleach/aliveBeforeBleach),
                other = (1-bleachSurvival)) %>% 
  select(`Outplant Date`, 
         `Site ID`, 
         species,
         Deliverable, 
         bleachSurvival,
         other) %>% 
  mutate_at(c("Site ID", 
        "Deliverable", "species"), as.factor) %>% 
  droplevels() %>% 
  mutate(`Outplant Date` = as.Date(`Outplant Date`, "%m %d %Y")) %>% 
  pivot_longer(cols = c("bleachSurvival", "other"), names_to = "monitoringCategory", values_to = "proportion") %>% 
  select(`Outplant Date`, `Site ID`,species, Deliverable, monitoringCategory, proportion) %>%
  mutate_at(c("monitoringCategory"), as.factor) %>%
  mutate(`Site ID` = factor(`Site ID`, levels = mixedsort(levels(`Site ID`))))

totalSurvival <- rbind(bleachSurvival1mo, bleachSurvival12mo) 




```
```{r}
myColors1 <- c("green3", "red3")



bleachSurvivalPlot2<- ggplot(totalSurvival, aes(x = `Site ID`, y = proportion, fill = monitoringCategory, color = I('black')))+
  geom_bar(position = "stack", stat = 'identity', width = 1)+
  ggtitle(expression("Post-Bleaching Event Proportional Survival"))+
  xlab("Site")+
  ylab("Proportional Survival")+
  scale_fill_manual(values = myColors1, labels = c("Alive", "Dead or Missing")) +
  facet_wrap(species ~ Deliverable, scales = "free_x")
  

bleachSurvivalPlot3 <- bleachSurvivalPlot2 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 40, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 40, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 30, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_blank(),
    legend.text = element_text(size = 30, face = "bold", colour = "black"),
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"),
    # Text above facets
    strip.text = element_text(size = 20))




ggsave("../figures/postBleachingSurvivalNEW.png", plot =bleachSurvivalPlot3, width = 30, height = 30, units = 'in', dpi = 600)

```

```{r}
bleachSurvival1moTab <- bd %>% 
  dplyr::filter(!`12-month Monitoring Date` %in% NA) %>% 
  dplyr::mutate(aliveBeforeBleach = (`total # frags outplanted`* `1-month % Survival`),
                deadAfterBleach = (`total # frags outplanted` * `2024 post-bleaching % dead`),
                aliveAfterBleach = (`total # frags outplanted` * `2024 post-bleaching % survival`),
                # aliveBeforeBleach = as.integer(aliveBeforeBleach),
                # aliveAfterBleach = as.integer(aliveAfterBleach),
                missingAfterBleach = (aliveBeforeBleach - deadAfterBleach+aliveAfterBleach),
                bleachSurvival = (aliveAfterBleach/aliveBeforeBleach),
                bleachDead = (deadAfterBleach/aliveBeforeBleach),
                bleachMissing = (missingAfterBleach/aliveBeforeBleach),
                other = (1-bleachSurvival)) %>% 
  select(`Outplant Date`, 
         `Site ID`, 
         species,
         Deliverable, 
         bleachSurvival,
         other,
         aliveBeforeBleach,
         aliveAfterBleach) %>% 
  mutate_at(c("Site ID", 
        "Deliverable", "species"), as.factor) %>% 
  droplevels() %>% 
  mutate(`Outplant Date` = as.Date(`Outplant Date`, "%m %d %Y")) %>% 
  mutate(`Site ID` = factor(`Site ID`, levels = mixedsort(levels(`Site ID`))))





bleachSurvival12moTab <- bd %>% 
  dplyr::filter(!`12-month Monitoring Date` %in% NA) %>% 
  dplyr::mutate(aliveBeforeBleach = (`total # frags outplanted`* `12-month % Survival`),
                deadAfterBleach = (`total # frags outplanted` * `2024 post-bleaching % dead`),
                aliveAfterBleach = (`total # frags outplanted` * `2024 post-bleaching % survival`),
                # aliveBeforeBleach = as.integer(aliveBeforeBleach),
                # aliveAfterBleach = as.integer(aliveAfterBleach),
                missingAfterBleach = (aliveBeforeBleach - deadAfterBleach+aliveAfterBleach),
                bleachSurvival = (aliveAfterBleach/aliveBeforeBleach),
                bleachDead = (deadAfterBleach/aliveBeforeBleach),
                bleachMissing = (missingAfterBleach/aliveBeforeBleach),
                other = (1-bleachSurvival)) %>% 
  select(`Outplant Date`, 
         `Site ID`, 
         species,
         Deliverable, 
         bleachSurvival,
         other,
         aliveBeforeBleach,
         aliveAfterBleach) %>% 
  mutate_at(c("Site ID", 
        "Deliverable", "species"), as.factor) %>% 
  droplevels() %>% 
  mutate(`Outplant Date` = as.Date(`Outplant Date`, "%m %d %Y")) %>% 
  mutate(`Site ID` = factor(`Site ID`, levels = mixedsort(levels(`Site ID`))))

totalSurvivalTab1 <- rbind(bleachSurvival1moTab, bleachSurvival12moTab) 

totalSurvivalTab <- totalSurvivalTab1 %>% 
  group_by(species) %>% 
  summarize(totalPreBleach = sum(aliveBeforeBleach),
                   totalRemaining = sum(aliveAfterBleach),
                   totalDeadOrMissing = (1-totalRemaining),
                   proportionSurived = (aliveAfterBleach/aliveBeforeBleach),
                   proportionDeadOrmissing = (totalDeadOrMissing/aliveBeforeBleach))
  




```


